Trafficking synaptic materials in complex morphologies involves tradeoffs in speed and precision

### 2015-08-21 latex/paper.tex
first pass at results done...ish

\author[1,2,3]{Alex H. Williams}
\author[2]{Cian O'Donnell}
\author[4]{Eve Marder}
\author[2,5]{Terrence Sejnowski}
\author[4]{Timothy O'Leary}

\affil[1]{Department of Neurosciences, University of California, San Diego, La Jolla, CA 92093, USA}
\affil[2]{Howard Hughes Medical Institute, Salk Institute for Biological Studies, La Jolla, CA 92037, USA}
\affil[3]{Department of Neurobiology, Stanford University, Stanford, CA 94305, USA}
\affil[4]{Volen Center and Biology Department, Brandeis University, Waltham, MA 02454, USA}
\affil[5]{Division of Biological Sciences, University of California at San Diego, La Jolla, CA 92093, USA}

@@

We have thusfar considered models that produce a desired distribution of cargo to arbitrarily high precision. However, real neurons probably tolerate some sloppiness in synaptic trafficking (or even utilize it as a substrate for complex plasticity rules). Thus, we examined the consequences of relaxing the separation of time scales between transport and detachment. In general, this improves the speed at which cargoes are delivered to synapses, but decreases the accuracy or specificity of delivery.

To appreciate this, consider an experiment where distal synaptic inputs on the apical tuft are stimulated (Fig. 5A). If the average detachment rate constants are sufficiently slow, then, as before, cargo is delivered selectively to the stimulated region (Fig. 5A, left). If we increase the detachment rates by a uniformly scaling, some cargo ``leaks off'' the microtubule path on its way to the distal synapses (Fig. 5A, right). We refer to this as a ``non-specific'' delivery of cargo, since cargo is not selectively delivered to the stimulated sites. \hl{In this case, non-specificity might induce on-the-path synaptic plasticity; a possibility that has some experimental evidence} \citep{Han_2013}.

Tradeoff curves between the average detachment rate constant and the non-specificity of transport for this stimulation pattern are shown in figure 5B. Importantly, this tradeoff exists for both trafficking strategies we examined in figure 4. In this case, the uniform trafficking strategy (blue line in Fig. 5B; see Fig. 4D) outperforms the specific trafficking strategy (red line in Fig. 5B; see Fig. 4C) since the latter suffers from bottleneck in the proximal apical zone.

@@

@@

Other assumptions of the mass-action model - Second, the model assumes are many copies of $u$ in each compartment. This allows us to approximate the mass as a continuous variable, and the dynamics of the system as deterministic. Third, the number of molecular motors and capacity of the microtubules is not saturated by the level of $u$ in any compartment. This means that the rate of transport from compartment A to compartment B linearly increases with the amount of $u$ in compartment A.

>>>>

\author[1,2,3,*]{Alex H. Williams}
\author[2]{Cian O'Donnell}
\author[4]{Eve Marder}
\author[2,5]{Terrence Sejnowski}
\author[4]{Timothy O'Leary}

\affil[1]{Department of Neurosciences, University of California, San Diego, La Jolla, CA 92093, USA}
\affil[2]{Howard Hughes Medical Institute, Salk Institute for Biological Studies, La Jolla, CA 92037, USA}
\affil[3]{Department of Neurobiology, Stanford University, Stanford, CA 94305, USA}
\affil[4]{Volen Center and Biology Department, Brandeis University, Waltham, MA 02454, USA}
\affil[5]{Division of Biological Sciences, University of California at San Diego, La Jolla, CA 92093, USA}
\affil[*]{Address correspondence to: ahwillia@stanford.edu}

@@

We have thusfar considered models that produce a desired distribution of cargo to arbitrarily high precision.
However, real neurons probably tolerate some sloppiness in synaptic trafficking (or even utilize it as a substrate for complex plasticity rules).
Thus, we examined the consequences of relaxing the separation of time scales between transport and detachment.
In general, this improves the speed at which cargoes are delivered to synapses, but decreases the accuracy of delivery.

To appreciate this, consider an experiment where distal synaptic inputs on the apical tuft are stimulated (Fig. 5A).
If the average detachment rate constants are sufficiently slow, then, as before, cargo is delivered selectively to the stimulated region (Fig. 5A, left).
If we increase the detachment rates by a uniformly scaling, some cargo ``leaks off'' the microtubule path on its way to the distal synapses (Fig. 5A, right).
We refer to this as a ``non-specific'' delivery of cargo, since cargo is not selectively delivered to the stimulated sites.
\hl{For this stimulation pattern, non-specificity might induce on-the-path synaptic plasticity; a possibility that has some experimental evidence} \citep{Han_2013}.

Tradeoff curves between the average detachment rate constant and the non-specificity of transport for this stimulation pattern are shown in figure 5B.
Importantly, this tradeoff exists for both trafficking strategies we examined in figure 4 --- the selective transport strategy (see Fig 4D) and uniform transport strategy (see Fig. 4C).
The uniform trafficking strategy (Fig. 5B, blue line) outperforms the specific trafficking strategy (Fig. 5B, red line) since the latter suffers from bottleneck in the proximal apical zone.

@@

To examine these observations over a larger range of stimulation patterns and transport parameters, we plotted tradeoff curves between delivery precision and speed.
We first examined a cable with six evenly spaced delivery sites (same as top row of Fig. 6A-C).
As before, a hard tradeoff between specifity and delivery speed exists for the untuned model (blue line in Fig. 6D, same transport rate constants as Fig. 6A and 6B).
As the transport rate constants are tuned to increase anterograde movement, the optimal points along the tradeoff curve move to the left, representing faster transport times.
However, the tradeoff curves also become nonmonotonic: the error (non-specificity) initially decreases as the detachment rate decreases, but begins to increase after a certain well-matched point.
Points on the descending left branch of the curve represent cargo distributions with proximal bias (detachment is too fast); points on the ascending right branch correspond to distal bias (detachment is too slow).

We observed qualitatively similar tradeoff curves for randomly positioned synapses (Fig. 6E), and for randomizing the number and position of synapses (Fig. 6F). There is substantial variability in the tradeoff curves in both cases, but particularly when randomizing the number of synaptic inputs. Notably, the untuned transport model (blue curves) always converge to zero error as the detachment rate decreases. Thus, for this model, it appears that the only way to achieve very precise and flexible transport is to have a slow detachment rate.

@@

Other assumptions of the mass-action model - Second, the model assumes are many copies of $u$ in each compartment.This allows us to approximate the mass as a continuous variable, and the dynamics of the system as deterministic.Third, the number of molecular motors and capacity of the microtubules is not saturated by the level of $u$ in any compartment. This means that the rate of transport from compartment A to compartment B linearly increases with the amount of $u$ in compartment A.
